---
title: "Manipulating data with dplyr Package"
author: "Suraj Bhattarai"
date: "September 26, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r} 
library(dplyr) # loads dplyr library
data(iris) # loads iris dataset
head(iris) # peek first 6 rows of data
dim(iris) # gives the dimension of data, there are 150 rows and 5 columns
any(is.na(iris))# there are no NAs in the dataset 
tail(iris)
```

### Answer 1. There are 150 observations and 5 variables in the iris dataset.

```{r}
iris1 <- filter(iris, Species %in% c("versicolor", "virginica"), Sepal.Length > 6, Sepal.Width > 2.5)
head(iris1)
dim(iris1)
```
```{r}
iris11 <- filter(iris, Species == c("versicolor", "virginica"), Sepal.Length > 6, Sepal.Width > 2.5)
dim(iris11)

filter(iris, Species == "versicolor" | Species == "virginica", Sepal.Length > 6, Sepal.Width > 2.5)
```

### Answer 2. There are 56 observations in the new dataset iris1 containing only versicolor and virginica species of sepal length larger than 6cm and sepal width larger than 2.5 cm.

```{r}
iris2 <- select(iris1, c(Species, Sepal.Length, Sepal.Width))
head(iris2)
dim(iris2)
```

### Answer 3. There are 56 observations and 3 variables in the iris2 dataset.

```{r}
iris3 <- arrange(iris2, desc(Sepal.Length))
head(iris3)
```

### Answer 4. First six rows of iris3 dataset arranged by Sepal.Length in descending order is shown by head command.

```{r}
iris4 <- mutate(iris3, Sepal.Area = Sepal.Length * Sepal.Width)
head(iris4)
dim(iris4)
```

### Answer 5. iris4 dataset contains 56 observations and 4 variables. A new variable Sepal.Area was introduced by mutate function 

```{r}
iris5 <- summary(iris4)
print(iris5)
```

### Answer 6. iris5 dataset was created which shows mean of all the variables including sample numbers by using summary function 

```{r}
iris6G <- group_by(iris4, Species)
iris6 <- summarise(iris6G, count = n(), length = mean(Sepal.Length), width = mean(Sepal.Width))
print(iris6)  
```

### Answer 7. iris6 dataset shows the mean length and width of each species. Different species were grouped by using group_by function

```{r}
iris %>%
  filter(Species %in% c("versicolor", "virginica"), Sepal.Length >6, Sepal.Width > 2.5) %>%
  select(c(Species, Sepal.Length, Sepal.Width)) %>%
  arrange(desc(Sepal.Length)) %>%
  mutate(Sepal.Area = Sepal.Length * Sepal.Width) %>%
  group_by(Species) %>%
  summarize(count = n(), length = mean(Sepal.Length), width = mean(Sepal.Width))
```

### Answer 8. all the manipulations done in a single code chunk using pipe %>%   


  
