---
title: "ggplot2"
author: "Suraj Bhattarai"
date: "December 12, 2018"
output: html_document
---
## ggplot syntax

ggplot(diamonds, aes(x=carat,y=price)) + geom_point()

 
The basic concept of a ggplot graphic is to combine different elements into *layers*. Each layer of a ggplotgraphic must have a data set and aesthetic mappings

* **data**: for **ggplot()**, this must be a data frame!
* **aes**: a mapping from the data to the plot; basically the x and y-axes


ggplot(diamonds, aes(x=carat,y=price)) + geom_point()

Layers can also have: 

* a  geom, or a geometric object: defines the overall look of the layer -- is it bars, points, or lines?

* a  stat, or a statistical summary: how should the data be summarized (e.g., binning for histograms, or smoothing to draw regression lines, etc).
* a position: how to handle overlapping points 

When not specified, the defaults are used.

 
We can use  **geom_boxplot** to create boxplots when one variable is continuous and the other is a factor. 

```{r,}
library(ggplot2)
ggplot(diamonds, aes(x=cut,y=price)) + geom_boxplot()

```


We can control the aesthetics of each layer, e.g.  *colour*, *size*, *shape*, *alpha* (opacity) etc.

```{r}
ggplot(diamonds, aes(carat, price)) + geom_point(colour = "blue")
```


Changing the alpha level

```{r}
ggplot(diamonds, aes(x=carat,y=price)) + geom_point(alpha = 0.2)
```


Changing the point size

```{r}
ggplot(diamonds, aes(x=carat,y=price)) + geom_point(size = 0.2)

```


Changing the shape and the point size
```{r}
ggplot(diamonds, aes(x=carat,y=price)) + geom_point(shape = 2,size=0.4)
```

 
The real power of **ggplot** is its ability to combine layers
```{r}
ggplot(diamonds, aes(x=carat,y=price)) + geom_point(size = 0.2) +
  geom_smooth()
```


In some cases, it may be more useful to get separate plots for each category of the third variable, to understand conditional relationships; using a facet wrap

```{r}
ggplot(diamonds, aes(carat, price)) + geom_point() +
  facet_wrap(~color, ncol=4)
```


Alternatively, you can use **facet_grid**, which also allows more than 1 conditioning variable (tables of plots)
```{r}
ggplot(diamonds, aes(carat, price)) + geom_point() +
 facet_grid(~color, labeller=label_both)
```


There are actually many ways to get the same plot! The following commands will produce the same plot:

* ggplot(diamonds, aes(price, carat)) + geom_point()
* ggplot() + geom_point(aes(price, carat), diamonds)
* ggplot(diamonds) + geom_point(aes(price, carat))
* ggplot(diamonds, aes(price)) + geom_point(aes(y = carat))
* ggplot(diamonds, aes(y=carat)) + geom_point(aes(x = price))

```{r}
ggplot(diamonds) + geom_point(aes(price, carat))
```

We can summarize univariate distributions using **boxplots**

```{r}
ggplot(diamonds, aes(1, depth)) + geom_boxplot()
```


However, a *histogram* would be a better choice here
```{r}
ggplot(diamonds, aes(depth)) + geom_histogram()
```

Notice the difference in the **aes** call; **boxplot** is really designed for multiple categories!


Tthe *default options* in **histogram** may not be sensible, and you often need to adjust the **binwidth** and **xlim**

```{r}
ggplot(diamonds, aes(depth)) + geom_histogram(binwidth=0.2) + xlim(56,67)
```

A better use of **boxplot** is when we want to compare distributions of a quantitative variable across categories of a factor variable, as previously discussed

```{r}
ggplot(diamonds, aes(cut, depth)) + geom_boxplot()
```


We can also get multiple histograms, though we need to either display them separately (less useful when comparing)
```{r}
ggplot(diamonds, aes(depth)) + geom_histogram(binwidth = 0.2) + 
    facet_wrap(~cut) + xlim(56, 67)
```

Or, we can overlay the historgrams 
```{r}
ggplot(diamonds, aes(depth, fill=cut)) + 
    geom_histogram(binwidth=0.2) + xlim(56,67)
```